Literature Review

### 2019-02-23 Naive_bayes_notes_5.Rmd
update

---
title: "Naive Bayes Workflow"
author: "Dan Johnson"
date: "12/21/2018"
output: html_document
---

@@

##Summary
This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments. I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here). This is quoted from Hansen's original tutorial:
>The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

>An example of the conditional probability that will be computed is as follows:

$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$

>which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.

Lines ~30-205 are the baseline version of the workflow using single words. It correctly classifies comments in the top 4 sub-categories  ~79% of the time. While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.

Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.

Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.

To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In these subsequent iterations I will evaluate:

*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
*  Using n-grams rather than single words
*  Making pair-wise comparison versus comparing multiple groups simultaneously
*  Modifying the variables in the NB analysis itself.

Finally, there are several things still missing or needed for this workflow:

*  How to feed a text dataset in, and get a table of the predicted categories back out.
*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
*  How to randomize which comments are used for training, testing datasets.

@@

##Baseline Version 1.0 of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

@@

```{r}
#Remove numbers. 
comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
```

@@

```{r}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
```

```{r}
#Stemming the text data to strip the suffix from words.  
comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
```

@@

###Basic Tokenizer
This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column. Be careful not to confuse it with "TermDocumentMatrix", which transposes the matrix.

"DocumentTermMaxtrix" function also can perform all of the text prep above in one command.

```{r}
comments_corpus_1gram <- DocumentTermMatrix(comments_corpus_clean)
comments_corpus_1gram_tidy <- tidy(comments_corpus_1gram)
inspect(comments_corpus_1gram)
```

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. 
```{r}
.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
comments_dtm_train <- comments_corpus_1gram[1:7005, ]
comments_dtm_test <- comments_corpus_1gram[7006:9340, ]
```

Calculate overall correct prediction rate.
```{r}
(42+454+1186+185)/23.35
```

***
##Modified Naive Bayes (NB) Method Using N-grams
The tokenizer command block above only works for single words. The block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command. The only change needed is "#" in "(words(x), #)" below. For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
}

@@

```{r}
(34+214+1254+94)/23.35
```

***
***
####\

##Visualization
This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
```{r}
wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
ONE <- subset(comments_raw, type == "1_Basic")
TWO <- subset(comments_raw, type == "2_Writing")
THREE <- subset(comments_raw, type == "3_Technical")
FOUR <- subset(comments_raw, type == "4_Logic")

wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

Show the 5 most frequent words in the data:
```{r}
comment_sack <- TermDocumentMatrix(comments_corpus_clean)
comment_pack <- as.matrix(comment_sack)
comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
head(comment_hack, 10)
```

And the 5 most frequent words from each class:
```{r}
wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

>>>>

---
title: "Permutations of Naive Bayes"
author: "Dan Johnson"
date: "12/21/2018"
output: html_document
---

@@

Uni-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 78.6%
  Highest Discrimination: 91.4%, 3_Technical

Bi-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 73.6%
  Highest Discrimination: 82.7%, 3_Technical

Uni-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 
  Highest Discrimination:

Bi-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 
  Highest Discrimination:

@@

###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

@@

```{}
#Remove numbers. 
comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
```

@@

```{}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
```

```{}
#Stemming the text data to strip the suffix from words.  
comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
```

@@

##Modified Naive Bayes (NB) Method Using N-grams
Uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command. The only change needed is "#" in "(words(x), #)" below.
```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

@@

```{r}
(44+375+1116+183)/23.35
```

***
***

### 2019-02-23 docs/hiding/classifier_oldversion.Rmd
update

---
title: "Data Import and Initial Exploration"
author: "Dan Johnson"
date: "12/17/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```
#Overview
This page has three sections. **Part 1** describes the target dataset for this project. **Part 2** outlines the data pre-checks and import process. **Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.

>>>>

---
title: "Classifier Development"
author: "Dan Johnson"
date: "12/17/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```
#Overview
This page has three sections. **Part 1** describes the target dataset for this project. **Part 2** outlines the data pre-checks and import process. **Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.

### 2019-02-23 docs/hiding/reference_tables.Rmd
update

---
title: "reference_tables"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

@@

## Reference Tables

Labels in **comment.subject** are routinely recoded as:

>>>>

---
title: "Reference Tables"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

@@

#TA Comment Categories
Labels in **comment.subject** are routinely recoded as:

### 2019-02-23 hiding/Xclassifier_oldversion.Rmd
update

---
title: "Classifier Development"
author: "Dan Johnson"
date: "12/17/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```
#Overview
This page has three sections. **Part 1** describes the target dataset for this project. **Part 2** outlines the data pre-checks and import process. **Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.

@@

#Part 1. Initial Dataset
The development dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.

#Part 2. Data Pre-Checks
Prior to import, review the data table **in Excel**; check that:

1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"

Column Number|Column Name|Accepted Values or Format
------------|----------------------|-----------------------
1|unique.record|Sp18.00001 (semester and year, then order in dataset)
2|report.id|R_0fanqHFhaY2yLT7 or similar
3|sort|remnant of coding
4|report.title (empty)|string
5|student|Std_drVZ2x1W (format is Std\_nnn)
6|course|113, 114, or 214
7|ta |TA_bc8EVX09 (format is TA\_nnn)
8|lab|113: allocation, betta; 114: manduca, frog; 214: enzymes (old) photosynthesis, physarum, cyanophage (new)
9|tag|first, second
10|type.TA|submission, revision
11|grade.TA|A, B, C, F
12|grading.time|time in minutes
13|Rank|remnant of coding
14|hypothesis.ok|Yes, No
15|data.ok|Yes, No
16|citation.ok|Yes, No
17|interpretation.ok|Yes, No
18|organization.ok|Yes, No
19|techflaws.ok|Yes, No
20|writing.ok|Yes, No
21|comments.incorporated|Yes, Somewhat, No
22|ta.comment|text string
23|code.subject|values in Ref. Tables
24|code.structure|values in Ref. Tables
25|code.locus|values in Ref. Tables
26|code.scope|values in Ref. Tables
27|code.tone|values in Ref. Tables
28|code.notes|values in Ref. Tables

Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**

>>>>

---
title: "Classifier Development"
author: "Dan Johnson"
date: "12/17/2018"
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```
#Overview
This page has three sections. **Part 1** describes the target dataset for this project. **Part 2** outlines the data pre-checks and import process. **Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.

@@

### 2019-02-23 hiding/projectnotes.Rmd
update

---
title: "Project Summary"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

# Abstract

# Introduction
## Background
Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors. Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space. Our practical goal is to develop techniques for teaching technical writing that:

*  Can be implemented at large scale by less experienced teaching assistants;
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."

We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses. In addition we have all comments that a TA made on the individual reports, plus associated metadata:

*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?

We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features. These insights then inform how we teach technical writing and train graduate instructors.

## Context for This Project & Prior Work
Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

In Summer 2018, we began analyzing TAs' comments from student papers. We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses. Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent. In all, the final tabulated dataset included ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

The sorted contingency table provided us with a relative estimate of TA effort.

## The Central Problem of This Project
Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

# Literature Review
##General Approach to Initial Dataset
The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*

@@

# Methods

@@

## Data

@@

# Results

@@

# Discussion

@@

# Conclusion

# References

>>>>

---
title: "Project Notes"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

##General Approach to Initial Dataset
This excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*

@@

## How to create files in R, then push to GitHub as a complete repo
Instructions for doing it on this website. [https://cfss.uchicago.edu/git05.html]()

@@

@@

@@

@@

### 2019-02-23 litreview.Rmd
update

---
title: "Literature Review"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Literature base

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

>>>>

---
title: "Literature Review"
date: "2/21/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This page links out to sub-pages that describe various background topics. Follow the links for details.

Description: https://adanieljohnson.github.io/default_website/pagename.html

Technical Writing Training

Text Classification Strategies

####\
